### 2024-02-27 _manuscript/_tex/index.tex
add all big files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  number]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.},
  pdfauthor={Simon Oiry; Bede Ffinian Rowe Davies; Pierre Gernez; Ana I. Sousa; Philippe Rosa; Maria Laura Zoffoli; Guillaume Brunier; Laurent Barillé},
  pdfkeywords={Drone, Remote Sensing},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Discriminating Seagrasses From Green Macroalgae in European
Intertidal areas using high resolution multispectral drone imagery.}
\author[1]{Simon Oiry%
\corref{cor1}%
}
 \ead{oirysimon@gmail.com} 
\author[1]{Bede Ffinian Rowe Davies%
%
}

\author[1]{Pierre Gernez%
%
}

\author[2]{Ana I. Sousa%
%
}

\author[1]{Philippe Rosa%
%
}

\author[3]{Maria Laura Zoffoli%
%
}

\author[4]{Guillaume Brunier%
%
}

\author[1]{Laurent Barillé%
%
}

\affiliation[1]{organization={Nantes University},,postcodesep={}}
\affiliation[2]{organization={Aveiro University},,postcodesep={}}
\affiliation[3]{organization={CNR ISMAR},,postcodesep={}}
\affiliation[4]{organization={JeSaisPas},,postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
Coastal areas host seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. To address these challenges, drone images
acquired with multi-spectral sensors offer a promising solution. This
study focuses on using drones acquisitions for mapping intertidal
macrophytes, effectively discriminating between seagrass and green
macroalgae. Ten drone flights were conducted at two different altitudes
(12m and 120m) across diverse European habitats in France and Portugal.
Low altitude flights were used to train a Deep Learning classifier based
on Neural Networks to discrimintate among 5 intertidal vegetation
classes. Drone mapping demonstrated an overall accuracy of 94\% across
all the sites and images, covering a total area of 467 000 m². The model
exhibited an accuracy of 96.4\% in identifying seagrass.
\end{abstract}

\begin{keyword}
    Drone \sep 
    Remote Sensing
\end{keyword}
\end{frontmatter}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\section{Introduction}\label{introduction}

Coastal areas are vital hotshots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between land and ocean \citep{unsworth2022}. These meadows offer a
myriad of ecosystem services to humanity, including limitation of
coastline erosion, reducing the risk of eutrophication, carbon
sequestration, and oxygen production. They serve as vital habitats for a
diverse array of marine and terrestrial species, providing living,
breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to their proximity to human activities, seagrass meadows are directly
exposed to and impacted by anthropogenic pressures. Global regression
and fragmentation are currently observed due to diseases, disasters,
coastal urbanization, sea reclamation, as well as fishing activities,
dredging, sea level rise, coastal erosion, competition with alien
species, and reduction in water quality \citetext{\citealp[
]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp{duffy2019}}. While
improvements in water quality have been recently reported in European
sites, allowing an overall recovery of seagrass ecosystems at the local
scale, many other coastal waters worldwide are still subjected to strong
eutrophication processes \citetext{\citealp[
]{deSantos2019}; \citealp{Zoffoli2021}}. Coastal eutrophication has been
associated to anomalous accumulation of green macroalgae, the so-called
green tides. Green tides produce shade and suffication over seagrass
individuals, thus threatening the health of seagrass ecosystems
\citetext{\citealp[ ]{Duarte2002}; \citealp{wang2022}}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of global and regional
monitoring programs for systematically surveying different Essential
Oceanic Variable \citep{Miloslavich2018} as seagrass coverage and
composition; as well as Essential Biodiversity Variable
\citep{Pereira2013} such as seagrass taxonomic diversity; species
distribution, population abundance, and phenology. Monitoring programs
also prioritize the identification of threats to these ecosystems,
particularly during early stages, to facilitate effective mitigation
actions. Traditionally, these ecological parameters have been quantified
through in situ measurements, although this approach faces several
constraints over intertidal zones. Intertidal meadows are only partially
exposed during low tide and can be situated in difficult-to-reach
mudflats, potentially leading to inaccurate and limited estimations with
conventional sampling techniques \citep{nijland2019}. However, satellite
data have been proven effective in complenmenting in situ surveys,
allowing for the rapid and consitent retrieval of EOV's over extensive
seagrass meadows. \citetext{\citealp[ ]{Zoffoli2021}; \citealp[
]{xu2021}; \citealp[ ]{Traganos2018}; \citealp{coffer2023}}

Satellite remote sensing offers the advantage of acquiring large-scale
data in real-time but presents its inherent challenges. Free access
satellite data (e.g., Sentinel-2 and Landsat8/9) provide relatively low
spatial resolution data (10 - 30 m) across a limited number of spectral
bands. These characteristics can be a limitation to accurately
discriminating seagrass from others co-existing macrophytes over the
meadow. Chlorophyceae (Green Algae) and marine Magnoliopsida (Seagrass)
share the same pigment composition \citetext{\citealp[
]{ralph2002}; \citealp{Douay2022}}. As a result, their respective
spectral signatures can be considered similar by a non-expert observer
\citetext{\citealp[ ]{Davies2023}; \citealp{bannari2022}}. Recently,
using a hyperspectral library, \citep{Davies2023} demonstrated that the
spectral resolution of Sentinel-2, might be enough for the
discrimination between Magnoliopsida and Chlorophyceae. However, green
tide events occur at small spatial scales that are not observable using
satellite imagery \citep{tuya2013}, especially during the initial stage
of the event.

Remote sensing drone acquisitions are presented as a tool that can
potentially fill gaps left by satellite and in situ data. Drone can
cover large expanses while recording imagery at significantly higher
spatial resolutions than satellite (pixel size from cm to mm) and still
capturing data at multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping the spatial
distribution of species \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, many studies showcase their findings with
study case limited to a single flight, restricting the generalizability
of their application to other sites \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. This study aimed
to analyze the potential of a drone equipped with a multispectral sensor
for maping intertidal macrophytes, with a particular focus on
discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were
performed over soft-bottom intertidal areas along two European countries
(France and Portugal), covering a wide range of habitats, from
monospecific seagrass meadows to meadows mixed with green or red algae.
A deep learning algorithm was trained and validated for macrophyte
discrimination, emphasizing applicability across diverse sites without a
loss of accuracy in predictions.

\section{Material \& Methods}\label{material-methods}

\subsection{Study sites}\label{study-sites}

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connect to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : One is located in one of these
island (Arz island). The other one is located in the sourthern part of
the Gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is a semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. With up to 6m in tidal range
during spring tide, a 6 km² seagrass meadow can uncover at low tide. The
sites we surveyed (L'Epine and Barbatre, Figure~\ref{fig-map} B) are
located in the southernmost part of the bay, along the Noirmoutier
Island, and host a monospecific \emph{Zostera noltei} seagrass bed with
very little mixing with other macrophytes. Three sites have been survey
in the Ria de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The
extent of this lagoon is about 80 km² with many narrow channels, large
salt marshes and many mudflats that uncovers at low tide. It is
connected to the open sea through a single channel, with a tidal lag
between the North and the South of the lagoon. The southernmost site
(Gafanha) is a mudflat located in an affluent of the lagoon whereas the
two other sites are situated in the middle of the lagoon and are only
accessible by boat (Figure~\ref{fig-map} C). These Portuguese sites were
characterized by a more diverse intertidal area, where the seagrass
meadow could intermingle with red algae (Rhodophyceae), brown algae
(Phaeophyceae), or green algae (Chlorophyceae).

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-map}
\begin{figure}[H]

\centering{

\includegraphics[width=26.67in,height=\textheight]{./Figures/Fig1_Map_Drone_Sites.png}

}

\caption{\label{fig-map}Sites where the different drone flights were
made. A: Gulf of Morbihan, B: Bourngeuf Bay, C: Ria de Aveiro Lagoon.
Light green represent terrestrial areas whereas darkgreen represent
intertidal areas.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Drone Data}\label{drone-data}

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in 10 spectral bands, from blue
(444 nm) to near infrared (840 nm). An angle of 90° has been maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each image has been set. Images have been
calibrated in reflectance using a calibration panel reflective at
\textasciitilde50\% provided by the camera's manufaturer and a
Downwelling Light Sensor (DLS2) that was utilized to acquire irradiance
data during the flight. A Structure-from-Motion photogrammetry software
(Agisoft Metashape) has been used to process images to obtain
multispectral orthomosaics of each flight. The workflow for
orthomosaicking has remained consistent for every flight. First, tying
key points were detected inside of each image and between overlapping
images in order to obtain a sparse point cloud. This cloud was cleaned
using reprojection accuracy metric in order to remove noisy points. A
dense point cloud has been then produced using a structure from motion
algorithm. A surface interpolation of this dense point cloud has been
made to obtain a Digital Surface Model (DSM), used to reconstruct the
multispectral ortho-image. Across all the sites, flights were made at
two different altitudes : 12 m or/and 120 m. Low altitude drone flights
produce ortho-images with a very high spatial resolution (8 mm per
pixel), making it simple to visually distinguish between the various
types of vegetation. High altitude fights on the other hand allow to
cover large areas and produced images with a pixel size of 80 mm (Table
1).

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{0pt}

\renewcommand*{\arraystretch}{1.5}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{ccccc}
\caption{List of drone Flight, summarising the date, the altitude and the purpose
of each flight.}\tabularnewline

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Country}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Site}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Name}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Altitude}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Utility}}} \\

\ascline{1.5pt}{000000}{1-5}\endfirsthead

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Country}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Site}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Name}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Altitude}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Utility}}} \\

\ascline{1.5pt}{000000}{1-5}\endhead

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Arz\ Island}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{12m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Duer}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{12m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{\multirow[c]{-3}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Morbihan}}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Duer}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{666666}{2-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Barbâtre}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{\multirow[c]{-5}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{France}}}}} & \multicolumn{1}{>{}c}{\multirow[c]{-2}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Bourgneuf\ Bay}}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{L'Epine}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{000000}{1-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Inner\ Lagoon}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Mataducos}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Gafanha}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

\ascline{1.5pt}{BEBEBE}{3-5}

\multicolumn{1}{>{}c}{\multirow[c]{-4}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Portugal}}}}} & \multicolumn{1}{>{}c}{\multirow[c]{-4}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Aveiro\ Lagoon}}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Gafanha}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{12m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training}}} \\

\ascline{1.5pt}{000000}{1-5}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Field sampling}\label{field-sampling}

Before each flight, targets used as Ground Control Points (GCPs) were
dispersed and georeferenced with a Trimble © Geo XH 6000 differential
GPS (dGPS). GCPs were used to correct georeferencing imprecision of
orthomosaics with an horizontal and vertical accuracy of 10cm. dGPS was
also used to georeference quardrats of 0.25cm² used to assess the
presence or absence of 5 key classes of vegetation : Bacillariophyceae
(Diatoms forming large biofilm over the sediment during low tide),
Phaeophyceae (Brown algae), Magnoliopsida (Seagrass), Chlorophyceae
(Green algae) and Rhodophyceae (Red algae)
(Figure~\ref{fig-vegetation}). Picture of each quadrat were taken and
uploaded on the online on the Global Biodiversity Information Facility
(GBIF) platform \citep{BedeGbif}. Each photograph was also processed to
estimate the percent cover of each type of vegetation. Georeferenced
hyperspectral reflectance signatures of each vegetation class were also
sampled using an ASD FieldSpec HandHeld 2 spectroradiomete, which
acquires reflectance between 325 and 1075 nm, with 1 nm of spectral
resolution. The methodological scheme followed in this study is
presented in Figure~\ref{fig-workflow}.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-vegetation}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{Figures/Spectral_shapes_total.png}

}

\caption{\label{fig-vegetation}Class of vegetation used to train the CNN
model and their standardised spectral signature}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Intertidal vegetation
mapping}\label{intertidal-vegetation-mapping}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-workflow}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Figure3_workflow.png}

}

\caption{\label{fig-workflow}Schematic representation of the workflow.
Diamonds represent input or output data, and rectangles represent Python
processing algorithms. The overall workflow of this study is divided
into two distinct parts based on the spatial resolution of the drone
flights: high-resolution flights were utilized for training and
prediction of the CNN model, whereas low-resolution flights were solely
employed for prediction purposes.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The spectral similarities of reflectance signatures between intertidal
green macrophytes (Magnoliopsida and Chlorophyceae) make their
discrimination challenging using simple classification algorithms
(Figure~\ref{fig-vegetation} F). To overcome this challenge, a deep
learning classification method was developed, trained, validated, and
applied to each drone flight (Figure~\ref{fig-workflow}).

\subsubsection{Neural Network model
building}\label{neural-network-model-building}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{0pt}

\renewcommand*{\arraystretch}{1.5}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{ccc}
\caption{Vegetation Classes of the model and the number of pixels used to train
and validate each class}\tabularnewline

\ascline{1.5pt}{666666}{1-3}

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Class}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training\ Pixels}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation\ Pixels}}} \\

\ascline{1.5pt}{666666}{1-3}\endfirsthead

\ascline{1.5pt}{666666}{1-3}

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Class}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Training\ Pixels}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation\ Pixels}}} \\

\ascline{1.5pt}{666666}{1-3}\endhead

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Bacillariophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{4,475}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{3,371,920}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Chlorophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{17,140}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{6,258,737}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Magnoliopsida}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{221,065}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{69,079,189}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Phaeophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{169,936}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{18,481,141}}} \\

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Rhodophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{5,771}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\

\ascline{1.5pt}{666666}{1-3}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Convolutional Neural Network (CNN) model with 2 hidden
layers. The training pixels were categorized into 7 different classes,
representing the various habitats encountered at the different study
sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae,
Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were
used for training purposes because of their 8 mm spatial resolution
allowing to avoid spectral sub-pixel mixing and to accurately visually
differentiate various vegetation classes. More than 418,000 pixels at 8
mm resolution from the 3 training flights have been used to train the
model (Table 2). Twenty one variables were used by the model as
predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX
multispectral camera (ranging from 444 nm to 840 nm), the same 10
spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{cao2017}) and the NDVI
(Equation~\ref{eq-ndvi}). Standardisation of spectral bands is used to
eliminate the scaling differences between spectra and to limit the
effect of biomass on the shape of the spectra \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

\begin{equation}\phantomsection\label{eq-std}{
R_{i}^{*}(\lambda) = \frac{R_{i}(\lambda) - min(R_{i})}{max(R_{i})- min(R_{i})}
}\end{equation}

where \(R_{i}(\lambda)\) is the reflectance at the wavelength
\((\lambda)\) of each individual spectra \((i)\), \(min(R_{i})\), and
\(max(R_{i})\) are the minimum and maximum value of the spectra \((i)\)

\begin{equation}\phantomsection\label{eq-ndvi}{
NDVI = \frac{R(840nm)-R(668nm)}{R(840nm)+R(668nm)}
}\end{equation}

where \(R(840nm)\) is the reflectance at 840 nm and \(R(668nm)\) is the
reflectance at 668 nm.

\subsubsection{Validation}\label{validation}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The CNN model was applied to all the 9 flights at both 12 and 120 m of
altitude. In situ information on georeferenced class type and percent
cover collected during each flight has been used to assess the CNN
accuracy. These images were used to construct a validation dataset
indicating the presence or absence of each class. Additionally to the
quadrat-based validation dataset, polygons of each class were photo
interpreted in order to increase the number of pixels of the validation
dataset. A confusion matrix, along with precision metrics such as the
global accuracy, the sensitivity, the specificity, and the Kappa
coefficient, was generated for each of the study sites. Finally each
validation matrix has been merge to create a unique matrix of the model.
A total of 536,000 pixels has been used to globally validate the CNN
model.

\section{Results}\label{results}

\subsection{Classification}\label{classification}

A total of 9 prediction maps corresponding to the 9 drone flights were
obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanaha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the 5
vegetation classes on which the CNN model was trained, 4 were present on
this site. On this site, there is a mixture of Chlorophyceae and
Rhodophyceae over the seagrass meadow. This is also where
Bacillariophyceae is most abundant. Although the seagrass bed is solely
composed of \emph{Zostera noltei}, various colors can be observed: dark
green (indicating healthy beds with 100\% coverage) and whitish/brown
(indicating beds where leaves are dying). Regardless of the color, the
meadow is predicted as Magnoliopsida by the CNN model.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-GafLow}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Maps Pred/FigX-Gaf_Low_Pred.png}

}

\caption{\label{fig-GafLow}RGB orthomosaic (Left) and Prediction (Right)
of the low altitude flight of Gafanaha, Portugal. The total extent of
this flight is 3000m² with a resolution of 8 mm per pixel. Background
colors means intertidal area (Light Green) and land area (Light Grey).
The zoom covers an area equivalent to a 10-meter Sentinel-2 pixel size.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
present in the flight, but it has been masked in the prediction map due
to glinting and misclassification caused by the turbid shallow water.
Most of the intertidal area has been classified as Magnoliopsida by the
model, even though there is discoloration of seagrass blades. Only a few
pixels have been classified as Chlorophyceae at this scale. Furthermore,
the area that was classified as Bacillariophyceae in the low-altitude
flight remains mostly classified as such in the high-altitude flight,
but some pixels of it are now classified as Magnoliopsida. The patches
of Rhodophyceae are still well classified. \emph{Spartina sp.} was
present in the northern part of the flight, near the land, and has been
classified either as Magnoliopsida or as Phaeophyceae.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-GafHigh}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Maps Pred/FigX-Gaf_High_Pred.png}

}

\caption{\label{fig-GafHigh}RGB orthomosaic (Left) and Prediction
(Right) of the high altitude flight of Gafanaha, Portugal. The total
extent of this flight is about 1 km² with a resolution of 80 mm per
pixel. Background colors means intertidal area (Light Green), land area
(Light Grey) and water (Light Blue). The red triangle shows the extent
of the low altitude flight of Gafanha. The zoom covers an area
equivalent to a 10-meter Sentinel-2 pixel size.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The High altitude flight made in the inner lagoon of the Ria de Aveiro
is the largest of all flights, covering almost 1.5 km²
(Figure~\ref{fig-Boat}). On this site, only seagrass and red algae where
seen on the field. The classification follows the same pattern, with a
patchy Magnoliopsida meadow mixed on the eastern part of the flight with
Rhodophyceae. As shown in the zoom (Figure~\ref{fig-Boat}), the edges of
the meadow can be colonised by Chlorophyceae.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-Boat}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Maps Pred/FigX-Boat_Pred.png}

}

\caption{\label{fig-Boat}RGB orthomosaic (Top) and Prediction (Bottom)
of the flight made in the inner part of Ria de Aveiro Lagoon, Portugal.
The total extent of this flight is about 1.5 km² with a resolution of 80
mm per pixel. Background colors means intertidal area (Light Green),
land area (Light Grey) and water (Light Blue). The zoom covers an area
equivalent to a 10-meter Sentinel-2 pixel size.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike crossing the northern
part of the flight from west to east. Alongside this dike, brown algae
attached to rocks and stranded green algae could be found. Despite the
high mixture between Chlorophyceae and Magnoliopsida in this flight,
these two classes are well discriminated by the classifier.

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-Dike}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Maps Pred/FigX-Dike_Pred.png}

}

\caption{\label{fig-Dike}RGB orthomosaic (Top) and Prediction (Bottom)
of Northern part of Noirmoutier Island, France. The total extent of this
flight is about 28 000 m² with a resolution of 80 mm per pixel.
Background colors means intertidal area (Light Green) and land area
(Light Grey). The zoom covers an area equivalent to a 10-meter
Sentinel-2 pixel size.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Validation}\label{validation-1}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-Validation}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Validation/ConfusionMatrixGlobal.png}

}

\caption{\label{fig-Validation}A global confusion matrix on the left is
derived from validation data across each flight, while a mosaic of
confusion matrices from individual flights is presented on the right.
The labels inside the matrices indicate the balanced accuracy for each
class. The labels at the bottom of the matrices indicate the User's
accuracy for each class, and those on the right indicate the Producer's
Accuracy. The values adjacent to the names of each site represent the
proportion of total pixels from that site contributing to the overall
matrix. Grey lines within the mosaic indicate the absence of validation
data for the class at that site. The table at the bottom summarizes the
Sensitivity, Specificity, and Accuracy for each class and for the
overall model.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

A total amount of 536,000 pixels have been used to validate the Neural
Network classifier. The site with the least amount of validation data is
Kerdrean with a total of 5557 pixels whereas Marinha Lanzarote is the
site with the most amount of validation data with 159713 pixels. The
global accuracy of the model is 94.26\% and a Kappa coefficient of 0.92
(Figure~\ref{fig-Validation}). The least performing site is Gafanha High
with an accuracy of 75.45\% whereas Mataducos is the site with most
accurate prediction, with a global accuracy of 98.05\%. Overall
Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae are well
classified with a balanced accuracy of 1, 0.96, 0.96 and 0.91
respectively. Bacillariophyceae is the least performing class with an
accuracy of 0.72 mainly due to a confusion with Magnoliopsida and
Sediment.

\subsection{Variable importance}\label{variable-importance}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-VIP}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{Figures/VIP/Fig_VIP.png}

}

\caption{\label{fig-VIP}Variable Importance of the Neural Network
Classifier for each vegetation class. The bigger the slice, the more
important is the variable to predict accuratly this class. The bottom
right plot shows the drone standardised reflectance spectra of each
class.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

The computation of the importance variable of each class indicate which
wavelength is mandatory to accurately predict the class
(Figure~\ref{fig-VIP}). Bands at 444 nm, 717 nm and 842 nm of the
Micasense camera are important for none of the vegetation classes. The
band at 531 nm is the only important predictor for the classifier to
accuratly predict chlorophyceae. In fact, at this wavelength, the
Chlorophyceae spectra has the maximum reflectance value of all the
classes. 531 nm and 740 nm are the most important predictors to predict
Phaeophyceae, corresponding to the minimum reflectance value among all
the classes. Bands at 475 and 560 nm are the most important predictors
for Bacillariophyceae and Rhodophyceae, respectively. 4 predictors,
ranging from the Green (560 nm) to the RedEdge (705 nm) are important to
predict accurately magnoliopsida.

\subsection{Effect of the flight height on the
prediction}\label{effect-of-the-flight-height-on-the-prediction}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-upscaling}
\begin{figure}[H]

\centering{

\includegraphics[width=0.9\textwidth,height=\textheight]{Figures/Upscaling/density_vs_Proportion.png}

}

\caption{\label{fig-upscaling}Kernel density plot showing the proportion
of pixel well classified based on the percent cover of the class in high
altitude flight pixels of Gafanha, Portugal. Each subplot shows all the
pixels of the same classes on the hight altitude flight. Percent cover
of classes is retrieve using the result of the classification of the low
altitude flight of Gafanha, Portugal. The vertical dashed line shows the
0.85 probability of the model. Everything on the right of this line has
a probability higher then 0.85 and everuthing on the left of this line
has a probability lower.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

Figure~\ref{fig-upscaling} is showing the percent cover of a class
required to classify correctly the high altitude flight. When the
percent cover of the class is 100 \%, big pixels are well classified for
all the classes excepted for Bare Sediment, where it's well classified
80\% of the time. A vegetative percent cover of at least 80\% is need to
have all the big pixels well classified, at the exception of
Magnoliopsida that needs an higher percent cover (\textgreater90 \%) to
be well classified. Concerning the probability of each class, a really
high Percent cover is needed to confidently predict Bacillariophyceae.
To predict Chlorophyceae with a confidence of 0.85, a percent cover of
93 \% is needed, 90 \% for magnoliopsida, 92 \% for Rhodophyceae and 97
\% for Bacillariophyceae.

\section{Discussion}\label{discussion}

\subsection{Seagrass Discrimination}\label{seagrass-discrimination}

\subsection{Spatial and Spectral Resolution impact on the
prediction}\label{spatial-and-spectral-resolution-impact-on-the-prediction}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-Pigm}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Disc_Pigment_Table.png}

}

\caption{\label{fig-Pigm}Photosynthetic and carotenoid pigments present
(Green) or absent (Red) in each taxonomic class present in the Neural
Network Classifier, along with their absorption wavelength measured with
spectroradiometer. Chl b: chlorophyll b, Chl c: chlorophyll c, Fuco:
fucoxanthin, Zea: zeaxanthin, Diato: diatoxanthin, Diadino:
diadinoxanthin, Neo: neoxanthin.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-CompareRef}
\begin{figure}[H]

\centering{

\includegraphics[width=0.5\textwidth,height=\textheight]{./Figures/Compare_reflectance.png}

}

\caption{\label{fig-CompareRef}Comparison of reflectance retrieved from
both low-altitude and high-altitude flights over a common area. The red
dashed line represents a 1 to 1 relationship.}

\end{figure}%

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\subsection{Big picture}\label{big-picture}

\section{Conclusion}\label{conclusion}

\renewcommand\refname{Bibliography}
  \bibliography{library.bib}

\end{document}